---
title: "Social Economic Conditions and Characteristics of Canadian Families"
author: "Yuanzhe Yang, Zichen Gong, Yukun Gao, Yujie He"
date: "October 19, 2020"
output:
  pdf_document: default
  pdf_html: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Title of your Report
Social Economic Conditions and Characteristics of Canadian Families

# Name(s) of Author(s) 
Yuanzhe Yang, Zichen Gong, Yukun Gao, Yujie He
# Date
October 19, 2020

## Abstract
The more people in a given household, the more likely that the family would be sensitive to socioeconomic conditions. If the more children the larger the household, and the higher the age the more likely to have more family members, old spouses at the age of 50 and above are more sensitive to socioeconomic conditions as compared to young married couples. The aim of the study is to quantitatively evaluate the condition based on historical data on surveys conducted from families in Canada. All data used in the analysis was collected in year 2017. 

The socioeconomic conditions considered in the study were education, income per family, and occupation. After running the linear regression model on the chosen variables, age had strong statistical significance to the total number of children. On the other hand, household size and conditions variables also had high statistical significance to the total number of children in a household. Therefore, the main conclusion is that the higher the age the higher the sensitivity to socioeconomic conditions as well as household size.

## Introduction
Trickett et al. (1991), indicated that families get affected by their socioeconomic status. Some of the major areas include levels of stability, parenting mechanisms as well as the growth outcomes in children. Fundamentally, many family members living in poverty are becoming victims of violence. 

Moreover, Health (no date), has indicated that the quality of life are presumably influenced by socioeconomic conditions like job opportunities, level of income, safety in society, and communal supports. 

A recent study by De Tray (1977), illustrated that spouses who have married for a long time additionally are likely to have more children. As a result, spouses that have lasted longer in marriage are likely to have larger households in terms of children being the additional family members as compared to new couples. 

Based on these studies, more experience in marriage and time with family suggests an increase in the household size; therefore, old age spouses pay more attention to socioeconomic conditions than young families. However, in many cases, the expected impacts are adverse in large households with higher sensitivity to the socioeconomic conditions. The driving factor is the economic status for families. Poorer families are likely to experience adverse impacts from socioeconomic conditions as compared to wealthier ones. Nonetheless, the study mainly focuses on the relationship between the household size and the sensitivity to socioeconomic conditions. 

The aim of this study is to assess the validity of the statement in quantitative aspects by evaluating families living in Canada. This frame population was particularly picked due to the availability of public data for surveys relating to family economic status in the country. 

## Data

The Dataset was extracted from the Microdata Analysis and Subsetting with SDA online database hosted by the University of Toronto, Faculty of Arts and Science. Nonetheless, the code used to clean the raw data is as provided by Alexander and Caeteno (2020). The dataset constitutes of figures gathered from a survey based on family structures found in Canada. This information was sourced from respondents from Canadian families where the main condition was that each respondent had to be 15 years and above. As a result, the survey results also constitute of data about the social demographics of parents. Moreover, the data collection was based on stratified sampling where the involved personnel divided the data into different clusters and subgroups for the purpose of representing the larger population. 
The target population in this case was involved persons with at least 15years and above living in Canada. The main focus was the residents of the Yukon, Northwest Territories as well as Nunavut. The main condition is that they had to be fulltime residents from various institutions. 
The Frame was created through the use of two main components. The first one was telephone numbers that include both landline and cellular that are available to Statistics Canada based on different sources such as telephone companies and census of population. The second component was the address register that covers all dwellings found in the chosen ten provinces. This frames were efficient since 86% of the available phone numbers in the country are connected to the address register. 
The total number of variables considered from the survey after performing a data cleanup was 77. This dataset is good since it provides an overall empirical view of structure of Canadian families from a wide perspective of quantitative variables and opinions.  Some of the important features include evaluation of families based on gender ages and occupation which is an important contribution to the study. The main drawback was that it was difficult to determine the purpose of certain variables in the study in the raw data form as they were not properly cleaned up.


```{r}
library(janitor)
library(tidyverse)

d <- read_csv("gss.csv")

```

## Model
The data was analyzed through the linear regression model. This model was based on the following formula: 

Total_children = a + B1age_first_child + B2age + B3feelings_life + B4hh_size + B5conditions + e

In this case, the overall size of children are considered as the dependent variable. This is based on the fact that they are attributed to being the main contributor to the size of a household, while age, age_of_first_child, feelings_life, and conditions are independent variables. They are the main factors influencing the outcome encased in this study for family socioeconomic standards. It was possible to assess how the total number of children in a family is related to variables such as age of parents and age of the first child. Moreover, the establishment of the relationship between total children at home and the life expectancy of guardians is also one perspective worth concerning. The results for running this model are discussed below.


## Results
The first step in understanding the data was to run a descriptive statistics code for interpreting how each variable is related to each other and their common attributes. The results for the descriptive statistics are as follows:

```{r}
summary(d)
```
Figure 1: Descriptive Statistics

Overall, some aspects worth noting in figure one is that the eldest person considered in the survey was in his/her 80s. Another fact to mention in the analysis is that the highest number of children in a family was 7.The maximum household size at any given time in the sample was 6 people.

Before running the linear regression, we should build up the scatter plot exploring how valid the results would be after running the regression. The results were as follows:

```{r}
head(d)

plot(d$total_children, d$age, xlab = 'Total Children per Household', ylab = 'Years old with Total Children', main= 'Regression Total Children to Age')
```
Figure 2: Scatter Plot

Essentially, total children and age variables in the study demonstrated some correlation based on the types of trends they have depicted from the analysis. As noted in figure one representation, the higher the age the more the number of children per family. The tabulation shows that there is a steady increase in a linear association and they mainly concentrate in one place as compared to another. As a result, it was possible to run the regression model as the results would be viable based on the topic of study. The regression results were as follows:

```{r}
mod <- lm(total_children ~ age + age_first_child
           + feelings_life + hh_size + conditions,d)

summary(mod)
```
Figure 3: Regression Model Results

The results indicate there is a strong association between the variables. Based on the nature of the study, the first consideration is the type of association between age and total_children. The results shown in figure one shows that there was statistical significance at the 0.001 critical levels. This outcome indicates that the higher the age the more the age contributes towards total children in a family. The outcome is further supported by the t value of 21.204. Essentially, given that the larger the absolute value of test statistic, the smaller the p-value; therefore the probability to reject the null hypothesis is higher. Age plays a significant role in affecting the number of children in families. A similar outcome is shown for the conditions variable where the p-value was significant at 0.001 critical level and the t-value was 7.057. 

In terms of the R-squared value, the results indicate that 0.5079/ 50.79% variance of the total_children can be predicted from the variables age and conditions. The only control variable showing a negative association is the feelings_life with a p-value of 0.12. This indicates that life expectancy plays a minimal role in controlling sensitivity towards socioeconomic conditions with regards to the size of a household.

## Discussion
The number of years in marriage does not only indicate elder spouses but also the size of a household (Lerman, 2002). A common phenomenon in this case is that at early years of marriage couples are less sensitive to socioeconomic conditions such as income, education, and occupation. However, as the number of household members increases, it contributes more towards the awareness on living standard. In essence, a two-person household is likely to have minimal pressure towards meeting obligations as compared to a household of ten family members where the needs are more excessive. This analysis shows that there is a strong association between age, household size, and total population of children in a family. 

# Weaknesses

First, there was a limitation for this data set that would support the validity of study, which is to evaluate the household size in terms of satisfaction. In this way it would be possible to establish a study on whether household size also contributes to adverse effects in terms of socioeconomic conditions.    

# Next Steps

Based on the study outcome, analysis on how socioeconomic conditions and age is related to occupation types by family members living in large households should be concerned. In this case it is possible to establish whether experience in a given occupation would contribute towards the sensitivity towards socioeconomic conditions. 

# Github Link

https://github.com/Yunglao12/Class-Group-Assignment-R-statistics/blob/main/ProblemSet2-new(1).Rmd

## References
Alexander, R. and Caeteno, S. (2020) GSS Code Cleaning 2017, University of Toronto. Available at: rohan.alexander@utoronto.ca.

Che, J. and Chen, J. (2001) ‘Food insecurity in Canadian households.’, Health reports / Statistics Canada, Canadian Centre for Health Information = Rapports sur la santé / Statistique Canada, Centre canadien d’information sur la santé.

Health, C. (no date) Social and Economic Factors | County Health Rankings & Roadmaps. Available at: https://www.countyhealthrankings.org/explore-health-rankings/measures-data-sources/county-health-rankings-model/health-factors/social-and-economic-factors (Accessed: 19 October 2020).

Lerman, R. I. (2002) ‘Marriage and the Economic Well-Being of Families with Children: A Review of the Literature’, (00).

De Tray, D. N. (1977) ‘Age of marriage and fertility: a policy review.’, Pakistan development review, 16(1), pp. 89–100. doi: 10.30541/v16i1pp.89-100.

Trickett, P. K. et al. (1991) ‘Relationship of Socioeconomic Status to the Etiology and Developmental Sequelae of Physical Child Abuse’, Developmental Psychology. doi: 10.1037/0012-1649.27.1.148.








